---
title: "making the network"
output: word_document
---
Remaking "adj" (copied code from similarity network.R).
```{r, eval=FALSE}
setwd("~/Box Sync/senior fall/Math Thesis")
courses <- read.csv("newCourses.csv", as.is=T)
courses <- courses[,-1]

n <- nrow(courses)
adj <- matrix (NA, nrow=n, ncol=n)
rownames(adj) <- courses$Dept1
colnames(adj) <- courses$Dept1

for(row in 1:(n-1)){
  rowList <- unlist(strsplit(courses$Keywords[row], " "))
  for(col in (row+1):n){
    if(row == col) next
     colList <- unlist(strsplit(courses$Keywords[col], " "))
     if((length(intersect(rowList, colList)) / length(union(rowList, colList))) == 1){
       adj[row, col] <- NA
       adj[col, row] <- NA
     } else{
       adj[row, col] <- length(intersect(rowList, colList)) / length(union(rowList, colList))
       adj[col, row] <- length(intersect(rowList, colList)) / length(union(rowList, colList))
     }
     
  }
}

```


Looking at isolated nodes - who are they?
```{r, eval=FALSE}
Adj <- adj >= 0.08
net <- graph_from_adjacency_matrix(Adj, mode="undirected")

which(degree(net) == 0)
which(degree(net) == 1)

sum(degree(net)==0)
sum(degree(net)==1)

## and super connected!
which(degree(net) > 50) #MORE LANGUAGE CLASSES!

```

Use adj to add edges (to those with degree 0 or 1) so all nodes are connected to at least two others (ensures game is playable & always possible to win).
```{r, eval=FALSE}
Adj <- adj >= 0.08
net <- graph_from_adjacency_matrix(Adj, mode="undirected")

## looking at network before
l <- layout_with_fr(net)
plot(net, layout=l,
     vertex.size = 1,
     vertex.label = "",
     main = "Network Before Connecting Low Degree Nodes")

###########

lowDegree <- which(degree(net) <= 1)

for(i in 1:length(lowDegree)){
  vertexName <- names(lowDegree[i])
  scores <- adj[vertexName,]
  
  ## if degree = 0
  if(degree(net, v = vertexName) == 0){
    edges <- names(sort(scores, decreasing=T)[1:2])
    net <- add_edges(net, c(vertexName, edges[1]))
    net <- add_edges(net, c(vertexName, edges[2]))
  }
  
  ## if degree = 1
  if(degree(net, v = vertexName) == 1){
    ## make sure no duplicate edges! check neighbors names 
    neighbor <- names(neighbors(net, v=vertexName))
    edges <- names(sort(scores, decreasing=T))
    if(edges[1] != neighbor){
      net <- add_edges(net, c(vertexName, edges[1]))
    }
    else {
      net <- add_edges(net, c(vertexName, edges[2]))
    }
  }
  
  ## if degree > 1 (bc already added edges)
  ## do nothing!
}

## looking at network after
l <- layout_with_fr(net)
plot(net, layout=l,
     vertex.size = 1,
     vertex.label = "",
     main = "Network After Connecting Low Degree Nodes")

max(degree(net))

```

This removes edges from the nodes with degree larger than 11 to make sure that all have degree <= 11 and connected AND min degree is still 2.
```{r, eval=FALSE}
## try with certain seed
testSeed <- 1

## finding large nodes
largeDegree <- names(which(degree(net) > 10))
neighbors(net, v = largeDegree[99])

## looping through and removing edges & checking connectivity
for(i in 1:length(largeDegree)){
  class <- largeDegree[i]
  counter <- 0 ## counter to make sure no dead ends
  
  while(length(neighbors(net,v = class)) > 11){
    counter <- counter + 1
    if(counter > 68){
      print("Completely stuck!")
    }
    testSeed <- testSeed + 1
    set.seed(testSeed)
    edgeName <- names(neighbors(net,v = class))[sample(length(neighbors(net,v = class)),1)]
    test <- delete_edges(net, paste(class,"|",edgeName, sep=""))
    
    ## checking connectivity & min degree
    if(is_connected(test) & min(degree(test)) == 2){
      ## ok to permanently remove edge
      net <- delete_edges(net, paste(class,"|",edgeName, sep=""))
    }
  }
}

## checking to make sure algorithm worked!
# checking max degree
max(degree(net))

# checking min degree
min(degree(net))

# checking connectivity 
is_connected(net)

## looking at network after
plot(net,
     vertex.size = 1,
     vertex.label = "",
     main = "Network After Removing Excess Edges")


## writing new network
#write_graph(net, "finalNetwork.txt", format = "ncol")

```

