---
title: "reading DB"
output: word_document
---

Reading in the database.
```{r}
library(RMySQL)
mydb = dbConnect(MySQL(), 
                 user='ktebbe', 
                 password='wordspassword', 
                 dbname='words', 
                 host='words.cl2itmibl6gz.us-west-2.rds.amazonaws.com')

```

looking at completed vs games given up
```{r}
# library(igraph)
# setwd("~/Box Sync/senior fall/Math Thesis/Shiny smart")
# net <- read_graph("networkDist.txt", format = "ncol", directed=FALSE)
# tab <- dbReadTable(mydb, "network3")
# 
# # completedSessions <- tab[which(tab$choice == tab$target), "session_id"]
# # uncompletedSessions <- setdiff(completedSessions, unique(tab$session_id))
# # 
# # ## validity check - WTFFF
# # length(completedSessions) + length(uncompletedSessions) == length(unique(tab$session_id))
# 
# ## calculating direct path length (remove cycles)
# ## func w session id as input
# sub <- tab[which(tab$session_id == completedSessions[1]),]
# sub <- sub[-which(sub$suggestion == 1),]
# trueLength <- nrow(sub)
# cycles <- sub$choice[duplicated(sub$choice)]
# 
# ## & remove suggestions
# shortestPaths <- shortest.paths(net, v=V(net), to=V(net))
# shortLength <- nrow(sub)
# actualLength <- shortestPaths[sub[1,"source"], sub[1,"target"]]
# 
# degSource <- degree(net, sub[1,"source"])
# degTarget <- degree(net, sub[1,"target"])

```

This function removes cycles from a single session. 
```{r}
removeCycles <- function(sub){
  if(length(sub$choice[duplicated(sub$choice)]) > 0) {
    KEEPGOING <- TRUE
    index <- 1
    while(KEEPGOING){
      cycles <- sub$choice[duplicated(sub$choice)]
      if(sub$choice[index] %in% cycles){
        ## find rows with that cycle and slice from latest one
        last <- max(which(sub$choice == sub$choice[index]))
        sub <- sub[-seq(from = (index+1), to = last, by = 1),]
      }
      index <- index + 1
      if(index > nrow(sub)){
        KEEPGOING <- FALSE 
      }
    }
  }
  return(sub)
}
```

This function corrects for any use of the "back" button.
```{r}
removeBacks <- function(sub){
  ## if back == 1 -> remove it and row above
  while(length(which(sub$back == 1)) > 0){
    backs <- which(sub$back == 1)[1]
    sub <- sub[-((backs-1):backs),]
  }
  return(sub)
}
```

For each session, measure:
-degree of source
-degree of target
-complete/incomplete
-traveled path length (traveled by user)
-cycle-free path length
-actual path length (by network attributes)
-number of "backs"
-number of "cycles"


```{r, eval=F}
library(igraph)
library(RMySQL)
mydb = dbConnect(MySQL(), 
                 user='ktebbe', 
                 password='wordspassword', 
                 dbname='words', 
                 host='words.cl2itmibl6gz.us-west-2.rds.amazonaws.com')

setwd("~/Box Sync/senior fall/Math Thesis/Shiny smart")
netSmart <- read_graph("networkDist.txt", format = "ncol", directed=FALSE)
tabSmart <- dbReadTable(mydb, "network3")

setwd("~/Box Sync/senior fall/Math Thesis/Shiny random")
netRandom <- read_graph("networkRandom.txt", format = "ncol", directed=FALSE)
tabRandom <- dbReadTable(mydb, "network2")

setwd("~/Box Sync/senior fall/Math Thesis/Shiny")
netBase <- read_graph("network.txt", format = "ncol", directed=FALSE)
tabBase <- dbReadTable(mydb, "network1")
tabBase <- tabBase[-1,] # first row was a tester

#########################

analyzeTable <- function(net, tab){
  df <- as.data.frame(matrix(NA, nrow = 1, ncol = 11))
  colnames(df) <- c("session_id", "completed",
                    "sourceDegree", "sourceName",
                    "targetDegree", "targetName",
                    "traveledPath", "cyclefree", "realPath",
                    "backs", "duplicates")
  shortestPaths <- shortest.paths(net, v=V(net), to=V(net))
  deg <- degree(net)
  idList <- unique(tab$session_id)
  
  for(i in 1:length(idList)){
    id <- idList[i]
    sub <- tab[which(tab$session_id == id), ]
    sub <- sub[sub$suggestion != 1,]
    ## fixing if they go back to first screen
    bad_ones <- sub$choice == ""
    sub$choice[bad_ones] <- sub$`source`[bad_ones]
    
    df[i, "session_id"] <- id
    df[i, "backs"] <- length(which(sub$back == 1))
    df[i, "duplicates"] <- length(unique(sub$choice[duplicated(sub$choice)]))
    df[i, "realPath"] <- shortestPaths[sub[1,"source"], sub[1,"target"]]
    df[i, "sourceDegree"] <- deg[sub[1,"source"]]
    df[i, "sourceName"] <- sub[1,"source"]
    df[i, "targetDegree"] <- deg[sub[1,"target"]]
    df[i, "targetName"] <- sub[1,"target"]
    df[i, "completed"] <- as.numeric(any(sub$choice == sub$target))
    
    sub <- removeBacks(sub)
    df[i, "traveledPath"] <- nrow(sub)
    sub <- removeCycles(sub)
    df[i, "cyclefree"] <- nrow(sub)
  }
  return(df)
}

smartResults <- analyzeTable(netSmart, tabSmart)
randomResults <- analyzeTable(netRandom, tabRandom)
baseResults <- analyzeTable(netBase, tabBase)

nrow(baseResults)
nrow(randomResults)
nrow(smartResults)

```

Visualization of completed vs incomplete for each network - barplot!
```{r, eval=F}
propBase <- length(which(baseResults$completed==1))/nrow(baseResults)
propSmart <- length(which(smartResults$completed==1))/nrow(smartResults)
propRandom <- length(which(randomResults$completed==1))/nrow(randomResults)

# m <- matrix(c(propBase, 1-propBase, 
#               propRandom, 1-propRandom,
#               propSmart, 1-propSmart),
#               ncol=3, nrow=2)
m <- matrix(c(propBase, 
              propRandom, 
              propSmart),
              ncol=3, nrow=1)
colnames(m) <- c("Base", "Random", "Smart")
b <- barplot(m, main = "Completion Rates For Each Network",
        ylab = "Proportion of Games",
        ylim = c(0,1))
text(b, m[1,]-.04, paste(round(m[1,],2)),
     col = "white")

```

Looking at mean degree of source/target in completed/uncompleted games.
```{r, eval=F}
s <- matrix(NA, ncol = 2, nrow = 6)
colnames(s) <- c("Source Degree", "Target Degree")
rownames(s) <- c("CompletedBase", "UncompletedBase", 
                 "CompletedRandom", "UncompletedRandom",
                 "CompletedSmart", "UncompletedSmart")

s[1,] <- c(mean(baseResults$sourceDegree[baseResults$completed==1]),
           mean(baseResults$targetDegree[baseResults$completed==1]))
s[3,] <- c(mean(randomResults$sourceDegree[randomResults$completed==1]),
           mean(randomResults$targetDegree[randomResults$completed==1]))
s[5,] <- c(mean(smartResults$sourceDegree[smartResults$completed==1]),
           mean(smartResults$targetDegree[smartResults$completed==1]))

s[2,] <- c(mean(baseResults$sourceDegree[baseResults$completed==0]),
           mean(baseResults$targetDegree[baseResults$completed==0]))
s[4,] <- c(mean(randomResults$sourceDegree[randomResults$completed==0]),
           mean(randomResults$targetDegree[randomResults$completed==0]))
s[6,] <- c(mean(smartResults$sourceDegree[smartResults$completed==0]),
           mean(smartResults$targetDegree[smartResults$completed==0]))

newS <- matrix(rowMeans(s), nrow=2, ncol=3)
colnames(newS) <- c("Base Network", "Random Network", "Smart Network")
b <- barplot(newS, beside=T, main = "Game Completion by Mean Degree for Each Network",
             ylab = "Mean Degree of Source & Target Vertices",
             ylim = c(0,7), xlim = c(1,9))
legend(x = .75, y = 7.2, c("Completed", "Not Completed"), pch = 15,
       col = grey.colors(2), cex = .9, bty = 'n')
```

look at language classes as destination -- hard!!!!

(betweenness?) centrality of source/target to predict success?

As baseline, look at mean shortest path length in each network, separated by completed/not
```{r, eval=F}
avgPath <- function(results){
  comp <- results[results$completed == 1, ]
  uncomp <- results[results$completed == 0, ]
  return(c(mean(comp$realPath), mean(uncomp$realPath)))
}
t <- matrix(NA, ncol = 2, nrow = 3)
colnames(t) <- c("Completed", "Not Completed")
rownames(t) <- c("Base Network", "Random Network", "Smart Network")

t[1,] <- avgPath(baseResults)
t[2,] <- avgPath(randomResults)
t[3,] <- avgPath(smartResults)
t

```


For completed games, look at mean difference in cyclefree path & shortest path.
```{r, eval=F}
t <- matrix(NA, ncol = 2, nrow = 3)
rownames(t) <- c("Base Network", "Random Network", "Smart Network")
colnames(t) <- c("Without Cycles", "With Cycles")

compDiffnoCycles <- function(results){
  comp <- results[results$completed == 1, ]
  return(mean(comp$cyclefree - comp$realPath))
}
t[1,1] <- compDiffnoCycles(baseResults)
t[2,1] <- compDiffnoCycles(randomResults)
t[3,1] <- compDiffnoCycles(smartResults)

all <- rbind(baseResults, randomResults, smartResults)
compDiffnoCycles(all)
compAll <-all[all$completed == 1, ]
uncompAll <-all[all$completed == 0, ]
summary(compAll$traveledPath)
summary(uncompAll$traveledPath)

compDiffwithCycles <- function(results){
  comp <- results[results$completed == 1, ]
  return(mean(comp$traveledPath - comp$realPath))
}
t[1,2] <- compDiffwithCycles(baseResults)
t[2,2] <- compDiffwithCycles(randomResults)
t[3,2] <- compDiffwithCycles(smartResults)
t

```

Looking at shortest path lengths across networks, separated by completion
```{r, eval=F}
shortestPathsBase <- shortest.paths(netBase, v=V(netBase), to=V(netBase))
#hist(shortestPathsBase, breaks = 10, xlim = c(0,10))

shortestPathsRandom <- shortest.paths(netRandom, v=V(netRandom), to=V(netRandom))
#hist(shortestPathsRandom, breaks = 10, xlim = c(0,10))

shortestPathsSmart <- shortest.paths(netSmart, v=V(netSmart), to=V(netSmart))
#hist(shortestPathsSmart, breaks = 10, xlim = c(0,10))

tab <- matrix(NA, nrow = 3, ncol = 3)
rownames(tab) <- c("Base Network", "Random Network", "Smart Network")
colnames(tab) <- c("Mean Path Length", "Diameter", "Mean Degree")

tab[,1] <- c(mean(shortestPathsBase), mean(shortestPathsRandom), mean(shortestPathsSmart))
tab[,2] <- c(max(shortestPathsBase), max(shortestPathsRandom), max(shortestPathsSmart))
tab[,3] <- c(mean(degree(netBase)), mean(degree(netRandom)), mean(degree(netSmart)))
tab
```

Looking at mean traveled path vs completion - how far do people go before giving up?
People who won went longer! except in smart network
```{r, eval=F}
avgTravel <- function(results){
  comp <- results[results$completed == 1, ]
  uncomp <- results[results$completed == 0, ]
  return(c(mean(comp$traveledPath), mean(uncomp$traveledPath)))
}

t <- matrix(NA, ncol = 2, nrow = 3)
colnames(t) <- c("Completed", "Not Completed")
rownames(t) <- c("Base Network", "Random Network", "Smart Network")

t[1,] <- avgTravel(baseResults)
t[2,] <- avgTravel(randomResults)
t[3,] <- avgTravel(smartResults)
t

```

Looking at different centrality metrics of the source & target nodes
```{r, eval=F}
eCent <- eigen_centrality(netSmart)
eCent["PSYC162"]

bCent <- betweenness(netSmart)
bCent["PSYC162"]


```

Looking at betweenness centrality scores of source/target in all, then split by completion

```{r, eval=F}
t <- matrix(NA, ncol = 3, nrow = 3)
colnames(t) <- c("All", "Completed", "Not Completed")
rownames(t) <- c("Base Network", "Random Network", "Smart Network")

avgBCent <- function(results, net){
  bCent <- betweenness(net)
  comp <- results[results$completed == 1, ]
  uncomp <- results[results$completed == 0, ]
  
  bAll <- mean(mean(bCent[results$sourceName]), mean(bCent[results$targetName]))
  bComp <- mean(mean(bCent[comp$sourceName]), mean(bCent[comp$targetName]))
  bUncomp <- mean(mean(bCent[uncomp$sourceName]), mean(bCent[uncomp$targetName]))
  return(c(bAll, bComp, bUncomp))
}

t[1,] <- avgBCent(baseResults, netBase)
t[2,] <- avgBCent(randomResults, netRandom)
t[3,] <- avgBCent(smartResults, netSmart)

t
## double WTF
```


```{r, eval=F}
mean(baseResults$duplicates)
mean(randomResults$duplicates)
mean(smartResults$duplicates)
```




