---
title: "similarity network"
output: word_document
---
```{r echo=FALSE, eval=FALSE}
setwd("~/Box Sync/senior fall/Math Thesis")
courses <- read.csv("courses.csv", as.is=TRUE)
courses <- courses[,-1]
```

We want to remove senior-only major classes. We will do so by creating a list of words, of which two appear in these "senior" classes, and removing them from the courses.csv file. 
```{r echo=FALSE, eval=FALSE}
## looking at words with two words from "senior" list
#courses <- read.csv("courses.csv", as.is=TRUE)
pairWords <- c("independent", "study", "senior", "project", "projects", "individual", "studies", "research", "special", "reading", "essay", "directed", "seminar", "topics", "colloquium", "junior", "tutorial")

titles <- courses$Title
titlesToRemove <- c()
for(i in 1:length(titles)){
  titleWords <- unlist(strsplit(titles[i], " ")) # split title into words
  titleWords <- tolower(titleWords) # make lower case
  if(length(intersect(titleWords, pairWords)) >= 2){
    titlesToRemove <- c(titlesToRemove, titles[i])
  }
}

## removing the courses if they match
for(i in 1:length(titlesToRemove)){
  row <- which(courses$Title == titlesToRemove[i])
  if(length(row) > 0) courses <- courses[-row,]
}

```

This code looks at most common words in all descriptions. This was used to create a list of words for the Shiny application, where people voted on the word's potency to illuminate what the course (that contains said word) is about. 
```{r echo=FALSE, eval=FALSE}
words <- c()
for(i in 1:nrow(courses)){
  words <- c(words, strsplit(courses$Description[i], " "))
}
words <- unlist(words,recursive=FALSE) # make one big list
words <- gsub("\\<[^\\]]*\\>", "", words, perl=TRUE) # removing anything between < >
words <- gsub("\n","",words) # removing \n notation

words <- unlist(strsplit(words, "&#160;"), recursive=FALSE) # splitting words by weird &#160; notation (also removes it)
words <- words[-grep("http:", words)] # removing anything with http://

words <- gsub("(?!-)[[:punct:]]", "", words, perl=TRUE) # removing punctuation except hyphens
words <- tolower(words) #make lower case

# remove stopwords
stop <- read.csv("stopwords.csv", as.is=T)
for(i in 1:nrow(stop)) {
  rows <- which(words == stop[i,1])
  if(length(rows) == 0) next
  words <- words[-rows]
}
#sort(table(words),decreasing=TRUE)[1:1000]
```

This creates the keyword list for each course, after removing the generic stop words and the words voted on from the Shiny application. 
```{r, eval=FALSE}
stopWords<- read.csv("stopwords.csv", as.is=T)
toRemove <- read.csv("wordsToRemove.csv", as.is=T)
toRemove <- toRemove[,3]
toRemove <- unique(toRemove)

## function that takes description and returns keywords
keepKeywords <- function(descrip){
  words <- strsplit(descrip, " ")
  words <- unlist(words,recursive=FALSE) # make one big list
  #splitting and removing weird string
  for(i in 1:length(words)){
    if(length(grep("&#160;",words[i])) > 0){
      words <- c(words,gsub( "&#160;.*$", "", words[i])) #before character
      words <- c(words,gsub( ".*&#160;", "", words[i])) #after character
      words <- words[-i] #string with character
    }
  }
  words <- gsub("\\<[^\\]]*\\>", "", words, perl=TRUE) # removing anything between < > -- merges words!! 636
  words <- gsub("\n","",words) # removing \n notation
  words <- gsub("[[:punct:]]","",words) # removing punctuation
  words <- tolower(words) #make lower case
  # removing stopwords
  for(i in 1:nrow(stopWords)) {
    rows <- which(words == stopWords[i,1])
    if(length(rows) == 0) next
    words <- words[-rows]
  }
  ## removing confimed bad words
  for(i in 1:length(toRemove)) {
    rows <- which(words == toRemove[i])
    if(length(rows) == 0) next
    words <- words[-rows]
  }
  
  ## remove duplicates
  words <- unique(words)

  # returning keywords, collapsed into single string - OR RETURN AS LIST??
  return(paste(words, collapse = ' '))
}

courses$Keywords <- NA
for(j in 1:nrow(courses)){
  courses$Keywords[j] <- keepKeywords(courses$Description[j])
}

```

Removing language courses that only contain the keywords "continuation", a department code (e.g. "SPAN"), and a course number. 
```{r looking at language courses, eval=FALSE}
rows <- c()
for(i in 1:nrow(courses)){
  if(length(unlist(strsplit(courses$Keywords[i], " ")))==3){
    rows <- c(rows, i)
  }
}
#length(rows)
#courses[rows,]
## taking out these rows!
courses <- courses[-rows, ]

#write.csv(courses, "newCourses.csv")
```

This chunk find the course with the most keywords - 57!
```{r echo=FALSE, eval=FALSE}
sapply(gregexpr("\\W+", courses$Keywords[1]), length) + 1 # count number of words in string

max <- 0
for(i in 1:nrow(courses)){
  words <- sapply(gregexpr("\\W+", courses$Keywords[i]), length) + 1 
  if(words > max){
    max <- words
    index <- i
  }
}
courses$Keywords[index]
```

This makes an adjacency matrix between all pairs of classes using the Jaccard index.  
```{r echo=FALSE, eval=FALSE}
n <- nrow(courses)
adj <- matrix (NA, nrow=n, ncol=n)
rownames(adj) <- courses$Dept1
colnames(adj) <- courses$Dept1

for(row in 1:(n-1)){
  rowList <- unlist(strsplit(courses$Keywords[row], " "))
  for(col in (row+1):n){
    if(row == col) next
     colList <- unlist(strsplit(courses$Keywords[col], " "))
     if((length(intersect(rowList, colList)) / length(union(rowList, colList))) == 1){
       adj[row, col] <- NA
       adj[col, row] <- NA
     } else{
       adj[row, col] <- length(intersect(rowList, colList)) / length(union(rowList, colList))
       adj[col, row] <- length(intersect(rowList, colList)) / length(union(rowList, colList))
     }
  }
}

```

This explores very similar courses and looks at their department codes and numbers. 
```{r echo=FALSE, eval=FALSE}
max(adj, na.rm=T)
mean(adj, na.rm=T) #very small!
median(adj, na.rm=T) #more similarities of 0 than anything else

#large <- which(adj >= .3 & adj >= .5, arr.ind = TRUE)
large <- which(adj >= .6, arr.ind = TRUE) 
sim <- as.data.frame(matrix(NA, ncol=4, nrow = nrow(large)))
colnames(sim) <- c("Dept1", "Dept1Title", "Dept2", "Dept2Title")
for(i in 1:nrow(large)){
  sim$Dept1[i] <- rownames(adj)[large[i,1]]
  sim$Dept2[i] <- colnames(adj)[large[i,2]]
  
  sim$Dept1Title[i] <- courses$Title[which(courses$Dept1 == rownames(adj)[large[i,1]])]
  sim$Dept2Title[i] <- courses$Title[which(courses$Dept1 == colnames(adj)[large[i,2]])]
}

#write.csv(sim, "highSim.csv")

hist(adj)
hist( adj[adj > 0.1] ,
      main = "Histogram of Similarity Scores (>0.1)",
      xlab = "Jaccard Index",
      xlim = c(0,1))
```

This looks at similarity scores within each department.
```{r echo=FALSE, eval=FALSE}
## pulling all unique dept codes
depts <- unique(gsub("[[:digit:]]", "", colnames(adj)))

withinScores <- matrix(NA, nrow = length(depts), ncol = 1)
colnames(withinScores) <- c("Score")
rownames(withinScores) <- depts

## looking at each unique dept
for(i in 1:length(depts)){
  matches <- adj[grep(depts[i], rownames(adj)), grep(depts[i], colnames(adj))]
  withinScores[i] <- mean(matches, na.rm=TRUE)
}

withinScores[which(is.nan(withinScores))] <- NA
```

This makes the betweenMax & betweenMean matrices.
```{r echo=FALSE, eval=FALSE}
# looking between departments
depts <- unique(gsub("[[:digit:]]", "", colnames(adj)))

betweenMax <- matrix(NA, nrow = length(depts), ncol = length(depts))
colnames(betweenMax) <- depts
rownames(betweenMax) <- depts

betweenMean <- matrix(NA, nrow = length(depts), ncol = length(depts))
colnames(betweenMean) <- depts
rownames(betweenMean) <- depts

## look at mean similarity between courses
for(i in 1:(length(depts)-1)){
  for(j in (i+1):length(depts)){
    matches <- adj[, grep(depts[i], colnames(adj))]
    if(is.null(dim(matches))){ 
      next }
    matches <- matches[grep(depts[j], rownames(matches)), ] ##
    if(length(matches) == 0){ 
      next }
    ## storing mean & max
    betweenMax[i,j] <- max(matches, na.rm = T)
    betweenMax[j,i] <- max(matches, na.rm = T)
    betweenMean[i,j] <- mean(matches, na.rm = T)
    betweenMean[j,i] <- mean(matches, na.rm = T)
  }
}

betweenMax <- as.data.frame(betweenMax)
betweenMean <- as.data.frame(betweenMean)

max(betweenMean, na.rm=T)

```

Let's look at different incremental cut-off values and network attributes.
```{r echo=FALSE, eval=FALSE}
library(igraph)
stats <- matrix(NA, nrow = 1, ncol = 4)
colnames(stats) <- c("Cutoff", "Isolated", "Max Degree", "Mean Degree")
stats <- as.data.frame(stats)
counter <- 1

for(i in 1:50){
  Adj <- adj >= (i * .01)
  net <- graph_from_adjacency_matrix(Adj, mode="undirected")
  stats[counter,] <- c(i*.01, 
                       sum(degree(net)==0),
                       max(degree(net)),
                       mean(degree(net)))
  counter <- counter + 1
  
}
stats
## logistic growth of isolated node number 
plot(stats$Cutoff,stats$Isolated, pch=16,
     main="Number of Isolated Nodes by Cutoff Score",
     xlab = "Jaccard Index Cutoff Score",
     ylab = "Number of Isolated Nodes")
lines(stats$Cutoff,stats$Isolated)

plot(stats$Cutoff,stats$`Mean Degree`, pch=16,
     main="Mean Degree by Cutoff Score",
     xlab = "Jaccard Index Cutoff Score",
     ylab = "Mean Degree of Network")
lines(stats$Cutoff,stats$`Mean Degree`)

plot(stats$Cutoff,stats$`Max Degree`, pch=16,
     main="Max Degree by Cutoff Score",
     xlab = "Jaccard Index Cutoff Score",
     ylab = "Max Degree in Network")
lines(stats$Cutoff,stats$`Max Degree`)

```

Same process, but focus in on a cutoff score of 0.05 - 0.08.
```{r echo=FALSE, eval=FALSE}
library(igraph)
stats <- matrix(NA, nrow = 1, ncol = 4)
colnames(stats) <- c("Cutoff", "Isolated", "Max Degree", "Mean Degree")
stats <- as.data.frame(stats)
counter <- 1

for(i in 1:30){
  val <- i * .001 + 0.049
  Adj <- adj >= val
  net <- graph_from_adjacency_matrix(Adj, mode="undirected")
  stats[counter,] <- c(val, 
                       sum(degree(net)==0),
                       max(degree(net)),
                       mean(degree(net)))
  counter <- counter + 1
  
}

## same plots!
plot(stats$Cutoff,stats$Isolated, pch=16,
     main="Number of Isolated Nodes by Cutoff Score",
     xlab = "Jaccard Index Cutoff Score",
     ylab = "Number of Isolated Nodes")
lines(stats$Cutoff,stats$Isolated)

plot(stats$Cutoff,stats$`Mean Degree`, pch=16,
     main="Mean Degree by Cutoff Score",
     xlab = "Jaccard Index Cutoff Score",
     ylab = "Mean Degree of Network")
lines(stats$Cutoff,stats$`Mean Degree`)

plot(stats$Cutoff,stats$`Max Degree`, pch=16,
     main="Max Degree by Cutoff Score",
     xlab = "Jaccard Index Cutoff Score",
     ylab = "Max Degree in Network")
lines(stats$Cutoff,stats$`Max Degree`)

Adj <- adj >= 0.05
net <- graph_from_adjacency_matrix(Adj, mode="undirected")
plot(net,
     vertex.size = 1,
     vertex.label = "")


```

